## Basic Programmable NSE

Non-Standard Evaluation (NSE hereafter) occurs when R expressions are
captured and evaluated in a manner different than if they had been executed
without intervention.  `subset` is a canonical example:

```{r}
subset(iris, Sepal.Width > 4.1)
```

`Sepal.Width` does not exist in the global environment, yet this works because
`subset` captures the expression and evaluates it within `iris`.

A limitation of NSE is that it is difficult to use programmatically:

```{r}
my.exp.a <- quote(Sepal.Width > 4.1)
subset(iris, my.exp.a)
```

`recsub` facilitates programmable NSE, as with this simplified version of
`subset`:

```{r}
subset2 <- function(x, subset) {
  sub.exp <- recsub(substitute(subset), x, parent.frame())
  sub.val <- eval(sub.exp, x, parent.frame())
  x[!is.na(sub.val) & sub.val, ]
}
subset2(iris, my.exp.a)
```

`recsub` is recursive:

```{r rec_ex_1}
my.exp.b <- quote(Species == 'virginica')
my.exp.c <- quote(Sepal.Width > 3.6)
my.exp.d <- quote(my.exp.b & my.exp.c)

subset2(iris, my.exp.d)
```

We abide by R semantics so that programmable NSE functions are almost
identical to normal NSE functions.  Using it them is the same, with
programmability as a bonus.  Recursive language substitution follows the
same semantics as normal object substitution.

