## Forwarding NSE Arguments to NSE Functions

If you wish to write a function that uses a programmable NSE function and
forwards its NSE arguments to it, you must ensure the NSE expressions are
evaluated in the correct environment, typically the `parent.frame()`.  This is
no different than with normal NSE functions.  An example:

```{r}
subset3 <- function(x, subset, select, drop=FALSE) {
  frm <- parent.frame()  # as per note in ?parent.frame, better to call here
  sub.q <- recsub(substitute(subset), x, frm)
  sel.q <- recsub(substitute(select), x, frm)
  eval(bquote(base::subset(.(x), .(sub.q), .(sel.q), drop=.(drop))), frm)
}
```

We use `bquote` to assemble our substituted call and `eval` to evaluate it in
the correct frame.  The parts of the call that should evaluate in `subset3` are
escaped with `.()`.  This requires some work from the programmer, but the user
reaps the benefits:

```{r}
col <- quote(Sepal.Length)
sub <- quote(Species == 'setosa')

subset3(iris, sub & col > 5.5, col:Petal.Length)
```

The forwarding is robust to unusual evaluation:

```{r}
col.a <- quote(I_dont_exist)
col.b <- quote(Sepal.Length)
sub.a <- quote(stop("all hell broke loose"))
sub.b <- quote(stop("all hell broke loose"))
threshold <- 3.35

local({
  col.a <- quote(Sepal.Width)
  sub.a <- quote(Species == 'virginica')
  subs <- list(sub.a, quote(Species == 'versicolor'))

  lapply(
    subs,
    function(x) subset3(iris, x & col.a > threshold, col.b:Petal.Length)
  )
})
```
