<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "README-",
  error = TRUE
)
library(recsub)
```

# recsub - Programmable Non-Standard Evaluation

[![](https://travis-ci.org/brodieG/recsub.svg?branch=master)](https://travis-ci.org/brodieG/recsub)
[![](https://codecov.io/github/brodieG/recsub/coverage.svg?branch=master)](https://codecov.io/github/brodieG/recsub?branch=master)
[![](http://www.r-pkg.org/badges/version/recsub)](https://cran.r-project.org/package=recsub)
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)

```{r child='vignettes/rmdhunks/basic.Rmd'}
```

## Additional Documentation

* [Intro vignette]() for a more in depth introduction to `recsub`.
* [NSE Functions with `recsub`]() in which we recreate simplified versions of
  `dplyr` and `data.table` that implement programmable NSE with `recsub`.

## Related Packages

[rlang](https://cran.r-project.org/package=rlang) by Lionel Henry and Hadley
Wickham implements a more comprehensive version of programmable NSE.  It is more
powerful, but also more complicated for the most common use of composing
NSE expressions to store in variables for later evaluation by an NSE function.

The following presentations are good resources for `rlang` in detail:

* [userR2017 Tidyeval Presentation][1]
* [Tidyeval and Hygienic Fexprs][2]

[1]: https://schd.ws/hosted_files/user2017/43/tidyeval-user.pdf
[2]: https://www.r-project.org/dsc/2017/slides/tidyeval-hygienic-fexprs.pdf

## Development Status

This package is proof-of-concept.  Feedback is welcome, particularly if you are
aware of some NSE pitfalls we may be ignoring.

Currently this package is only available on github:

```{r eval=FALSE}
devtools::install_github('brodieg/recsub')
```

## Acknowledgements

* R Core for developing and maintaining such a wonderful language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository, and Uwe Ligges in particular for maintaining
  [Winbuilder](http://win-builder.r-project.org/).
* [Jim Hester](https://github.com/jimhester) because
  [covr](https://cran.r-project.org/package=covr) rocks.
* [Dirk Eddelbuettel](https://github.com/eddelbuettel) and [Carl
  Boettiger](https://github.com/cboettig) for the
  [rocker](https://github.com/rocker-org/rocker) project, and [Gábor
  Csárdi](https://github.com/gaborcsardi) and the
  [R-consortium](https://www.r-consortium.org/) for
  [Rhub](https://github.com/r-hub), without which testing bugs on R-devel and
  other platforms would be a nightmare.
* Hadley Wickham for [devtools](https://cran.r-project.org/package=devtools) and
  [roxygen2](https://cran.r-project.org/package=roxygen2).
* [Yihui Xie](https://github.com/yihui) for
  [knitr](https://cran.r-project.org/package=knitr) and  [J.J.
  Allaire](https://github.com/jjallaire) etal for
  [rmarkdown](https://cran.r-project.org/package=rmarkdown), and by extension
  John MacFarlane for [pandoc](http://pandoc.org/).

