<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "README-",
  error = TRUE
)
library(matrioshka)
```

# matrioshka - Simple Programmable NSE

[![](https://travis-ci.org/brodieG/matrioshka.svg?branch=master)](https://travis-ci.org/brodieG/matrioshka)
[![](https://codecov.io/github/brodieG/matrioshka/coverage.svg?branch=master)](https://codecov.io/github/brodieG/matrioshka?branch=master)
[![](http://www.r-pkg.org/badges/version/matrioshka)](https://cran.r-project.org/package=matrioshka)
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)

```{r child='vignettes/rmdhunks/basic.Rmd'}
```

## Documentation

* [Intro
  vignette](http://htmlpreview.github.io/?https://github.com/brodieG/matrioshka/blob/master/inst/doc/introduction.html)
  for a more in depth introduction to `matrioshka`, including a brief comparison to
  `rlang`.
* [NSE Functions with
  `matrioshka`](http://htmlpreview.github.io/?https://github.com/brodieG/matrioshka/blob/master/inst/doc/nse-fun.html)
  in which we recreate simplified versions of `dplyr` and `data.table` that
  implement programmable NSE with `matrioshka::expand`.

## Installation

```{r eval=FALSE}
install.packages('matrioshka')
```

This package is proof-of-concept.  Feedback is welcome, particularly if you are
aware of some NSE pitfalls we may be ignoring.

## Acknowledgements

* R Core for developing and maintaining such a wonderful language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository, and Uwe Ligges in particular for maintaining
  [Winbuilder](http://win-builder.r-project.org/).
* [Jim Hester](https://github.com/jimhester) because
  [covr](https://cran.r-project.org/package=covr) rocks.
* [Dirk Eddelbuettel](https://github.com/eddelbuettel) and [Carl
  Boettiger](https://github.com/cboettig) for the
  [rocker](https://github.com/rocker-org/rocker) project, and [Gábor
  Csárdi](https://github.com/gaborcsardi) and the
  [R-consortium](https://www.r-consortium.org/) for
  [Rhub](https://github.com/r-hub), without which testing bugs on R-devel and
  other platforms would be a nightmare.
* Hadley Wickham for [devtools](https://cran.r-project.org/package=devtools) and
  [roxygen2](https://cran.r-project.org/package=roxygen2).
* [Yihui Xie](https://github.com/yihui) for
  [knitr](https://cran.r-project.org/package=knitr) and  [J.J.
  Allaire](https://github.com/jjallaire) etal for
  [rmarkdown](https://cran.r-project.org/package=rmarkdown), and by extension
  John MacFarlane for [pandoc](http://pandoc.org/).

## About

Brodie Gaslam is a hobbyist programmer based on the US East Coast.

The recursive substitution of symbols evokes Russian nesting dolls, hence
the name of the package.  We would have preferred to name the `expand` function
`unnest`, but the symbol is already in use in the popular `tidyr` package.

